A Survey of Fuzzing and New Technique for Dynamic Analysis

### 2014-02-23 docs/paper/fsuthesis/thesis-template/paper.tex
about 2/3 through with brumleys paper

Like TODO Owen's thing, Kudzu tries to circumvent the issue of solving complicated constraints by first converting JavaScript into an intermediate language called JASIL. Strings, in particular, are parsed and partitioned into several sub types which are more uniform. For instance, one type handles strings which are semantically regular expressions, another handles strings which are the concatenation of two other strings. Although not as featureful as JavaScript strings, this IL captures a large enough number of String features to allow Kudzu to find previous unknown exploits in programs. Shortcomings include an incomplete support of String operations and JavaScript regular expressions, notably backreferences, which match a previously matched string in the regex. Although representation of strings is incomplete, Kudzu reasons well enough to find two new vulnerabilities, as well as 9 previous vulnerabilities [AD].

@@

TODO BRUMLEY

DTA is the process of locating sources of taint and following their propagation through a binary to data sinks. Since, for exploitation, bugs must be triggered by user input, DTA usually identifies all user input (files, command line arguments, UI interactions, ...) as tainted, and marks it accordingly.

The granularity of taint marking is variable. Assigning everything from a generic binary value of tainted/not tainted to bit-level marking of which part of user input affected which variable is conceivable, although finer-grained taint analysis is more computationally expensive and difficult to determine.

@@

Handling taint correctly is of great important because of the existence of subtle implicit flow cases referenced by Clause, J. et al [D]. Implicit data flows are not always considered by dynamic tainting techniques [E], but for my project, implicit data flows are very important, so I need to implement DTA which catches both implicit and explicit flows.

DTA has been used to implement smart mutation fuzzers [A, T, MORE], but I think my work is unique from and complementary to Bekrar's work. In [A], DTA is proposed as a way to intelligently decide which parts of user input should be mutated. In [T], dynamic taint analysis identified hot bytes, which were then modified randomly or with boundary values.

I plan to add on to these ideas by allowing smarter mutations to be selected by backsolving to decide which values are most likely to result in new paths being explored. This has been done with symbolic execution before, but traditional symbolic execution is much more heavy-weight, both in time required to setup the environment and in overhead introduced to run the program, than my proposed approach.

>>>>

Similar to SimpIL [C], Kudzu tries to circumvent the issue of solving complicated constraints by first converting JavaScript into an intermediate language called JASIL. Strings, in particular, are parsed and partitioned into several sub types which are more uniform. For instance, one type handles strings which are semantically regular expressions, another handles strings which are the concatenation of two other strings. Although not as featureful as JavaScript strings, this IL captures a large enough number of String features to allow Kudzu to find previous unknown exploits in programs. Shortcomings include an incomplete support of String operations and JavaScript regular expressions, notably backreferences, which match a previously matched string in the regex. Although representation of strings is incomplete, Kudzu reasons well enough to find two new vulnerabilities, as well as 9 previous vulnerabilities [AD].

@@

In [C], the authors summarize dynamic taint analysis and symbolic execution and attempt to formalize each technique. A formal definition for DTA requires a formal language upon which to perform taint analysis, so the authors use SimpIL, a Simple Intermediate Language. Note that this is distinct from but related to the intermediate representations which some fuzzing research converts sources into. When fuzzers first convert to an intermediate language, this is done to simplify future analysis, as when Kudzu [AD] converts JavaScript strings into one of several new data types which are easier to reason about. Time is traded up front in order to make future operations easier or possible. SimpIL is a complete programming language, lacking in features but used as a base for some of the author's other work. It should be noted that the features SimpIL lacks, in particular functions and variable scoping, can be implemented using SimpIL, and SimpIL expressions not having side effects similarly do not hinder it. The important implementation issues for DTA spring from one's tainting policy. Our tainting policy decides under what circumstances we introduce taint, verify it, and track its propagation throughout the program execution [C].

Introduction rules should specify when taint first appears in a program, what type of taint it is, and what parts of the program have become tainted as a result. We can record taint at varying levels of granularity, with the tradeoff of finer-grained taint requiring greater system resources and slowing down analysis. Different types of taint may be more interesting to different kinds of fuzzers because different input comes from sources of varying trustedness [C]. For instance, we should probably trust raw, unsanitized user input we have read less than we trust a configuration file that only privileged users can modify. Both might be read in as input to a program, but the former is much more likely to contain malicious data like shellcode trying to exploit a vulnerability. Fuzzers have continued to drill down into finer and finer levels of granularity, from assigning everything with a generic binary value of tainted/not tainted to byte and even bit-level marking of which part of user input affected which variable. [CITE] [C] Note that this also expands taint's data type from a bit to something larger, and each tainted object requires this tag.

Just as with taint introduction, we verify taint based on our overall objective. If we are worried about a binary being exploited to execute arbitrary code, we may instrument the EIP register. When we verify that taint has reached EIP, we know we stand a strong chance of having found an exploitable condition. Likewise, were we monitoring taint not in order to generate test cases for fuzzing, but to study a piece of malware, we might instead watch for when a particular routine, which we think is unpacking the malware, has finished, and the malware is now available in an unobfuscated form.

Taint propagates through a program from its initial source, perhaps making its way to vulnerable sinks. The condition we are interested in is when user input finds its way to these vulnerable sinks in a program, and is able to influence them. As an example, user input received on STDIN which is able to change the value of the EIP register in x86 assembly is able to hijack control flow in the program. This may be as innocuous, though annoying, as forcing EIP to strange areas in the binary, resulting in segfaults, to more sinister remote code execution via pointing EIP to user controlled shellcode.

There are several challenges involved with implementing DTA. First of all, accurate taint analysis, as free as possible from both over and under tainting, requires a precisely implemented taint policy. Different types of control flows, either direct or indirect, must be handled as well, and one-way or lossy functions also tend to be either difficult or untenable to instrument.

TODO under/over

@@

Handling taint correctly is of great importance because of the existence of subtle implicit flow cases referenced by Clause, J. et al [D]. Implicit data flows are not always considered by dynamic tainting techniques [E], but for my project, implicit data flows are very important, so I need to implement DTA which catches both implicit and explicit flows.

TODO one-way

DTA has been used to implement smart mutation fuzzers [A, T, MORE], but I believe the technique I explore later is unique from and complementary to Bekrar's work. In [A], DTA is proposed as a way to intelligently decide which parts of user input should be mutated. In [T], dynamic taint analysis identified hot bytes, which were then modified randomly or with boundary values. I plan to add on to these ideas by allowing smarter mutations to be selected by backsolving to decide which values are most likely to result in new paths being explored. This has been done with symbolic execution before, but traditional symbolic execution is much more heavy-weight, both in time required to setup the environment and in overhead introduced to run the program, than my proposed approach.
